---
title: "UK Tax Receipts with hmrc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{UK Tax Receipts with hmrc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The `hmrc` package provides tidy access to statistical data published by HM Revenue
and Customs (HMRC) on GOV.UK. This vignette demonstrates the two core functions
covering monthly tax receipts and property transactions.

## Tax receipts

`get_tax_receipts()` downloads the monthly HMRC Tax Receipts and National Insurance
Contributions bulletin. The bulletin covers all major UK taxes and duties from
April 2016 to the most recent published month.

```{r receipts-basic}
library(hmrc)

# All tax heads — returns ~41 series
receipts <- get_tax_receipts()
head(receipts)
```

Filter to specific tax heads using the `tax` argument. Use `list_tax_heads()` to
see all available identifiers:

```{r list-heads}
list_tax_heads()
```

```{r receipts-filter}
# The three largest revenue streams
big_three <- get_tax_receipts(
  tax   = c("income_tax", "vat", "nics_total"),
  start = "2020-01"
)
```

```{r receipts-plot, fig.width=7, fig.height=4}
library(ggplot2)

ggplot(big_three, aes(x = date, y = receipts_gbp_m / 1000, colour = description)) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(labels = scales::label_comma(suffix = "bn")) +
  labs(
    title  = "UK monthly tax receipts",
    x      = NULL,
    y      = "GBP billions",
    colour = NULL,
    caption = "Source: HMRC Tax Receipts and NICs bulletin"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## Property transactions

`get_property_transactions()` downloads the monthly property transaction counts
for England, Scotland, Wales, Northern Ireland, and the UK total. Data runs from
April 2005.

```{r property-basic}
# UK residential transactions since 2018
prop <- get_property_transactions(
  type   = "residential",
  nation = "uk",
  start  = "2018-01"
)
head(prop)
```

```{r property-plot, fig.width=7, fig.height=4}
ggplot(prop, aes(x = date, y = transactions / 1000)) +
  geom_line(colour = "#3B82F6", linewidth = 0.8) +
  geom_smooth(method = "loess", se = FALSE, colour = "#94A3B8",
              linewidth = 0.6, linetype = "dashed") +
  scale_y_continuous(labels = scales::label_comma(suffix = "k")) +
  labs(
    title   = "UK residential property transactions",
    x       = NULL,
    y       = "Transactions (thousands)",
    caption = "Source: HMRC Monthly Property Transactions bulletin"
  ) +
  theme_minimal(base_size = 12)
```

## Caching

All downloads are cached locally. The cache persists between R sessions so
subsequent calls are instant. Use `clear_cache()` to delete cached files:

```{r cache}
# Remove files older than 30 days
clear_cache(max_age_days = 30)

# Remove everything
clear_cache()
```
